---
title: "Assignment_02"
author: "Fatima Arshad"
date: "2/9/2022"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
This case study is about predicting the rental price for an apartment offered by Airbnb in Vienna, Austria.The goal is to predict the price that may be appropriate for an apartment with certain features. This prediction is to help the company to answer the following business question: to predict prices for apartments in Vienna, Austria. We will take expected rental price for various kinds of apartment at various locations of the city into consideration.

The case study uses the airbnb dataset dated 08 December 2021. The dataset has a single data table that includes 4499 observations. The data refer to rental prices for one night in March 2017. The target variable is price per night per person, expressed in US dollars, because it was US dollars in the raw data. 

We are predicting price. Thus, we should ask whether we should leave the price variable as it is or take its log. The two histograms in Figure 14.3 show the distribution of price and log price.




```{r}
###########################     DATA CLEANING   ########################

#setting working directory
rm(list=ls())

library(tidyverse)

## GETTING THE DATA 
# download the data from https://github.com/fatimamalikk/data-analysis-3/tree/main/Assignment_02_Data/Raw
# then set your path to be able to run the code (I can't link it directly as it is too big, so you have to download)

path <- "C:/Users/4star/Documents/DA3/Git Assignments R/data-analysis-3/Assignment_02_Data"

# location folders
data_in  <- paste0(path,"/Raw/")
data_out <- paste0(path,"/Clean/")

# load the data
df <- read_csv(paste0(data_in,"Vienna_listings.csv"))


# INITIAL STEPS -----------------------------------------------------------

# basic data checks
sort(unique(df$last_scraped)) # they were scraped between "2021-12-08" "2021-12-09" "2021-12-18"
sum(rowSums(is.na(df)) == ncol(df)) # no only NA rows
nrow(df[duplicated(df),]) #  no duplicates 
sum(colSums(is.na(df)) > 0.5*nrow(df)) # there are 4 columns with at least 50% of the values missing 


# drop unnecessary columns
drops <- c("listing_url",
           "last_scraped",
           "name",
           "description",
           "neighborhood_overview",
           "picture_url",
           "host_url",
           "host_location",
           "host_about",
           "host_thumbnail_url",
           "host_picture_url",
           "host_total_listings_count",
           "neighbourhood_group_cleansed",  # only NA values in column
           "bathrooms", # only NA values in column
           "minimum_minimum_nights", "minimum_maximum_nights", "minimum_nights_avg_ntm",
           "maximum_minimum_nights", "maximum_maximum_nights", "maximum_nights_avg_ntm", 
           "calendar_updated", # only NA values in column
           "number_of_reviews_ltm", "number_of_reviews_l30d",
           "calculated_host_listings_count_entire_homes", 
           "calculated_host_listings_count_private_rooms", 
           "calculated_host_listings_count_shared_rooms")

df<-df[ , !(names(df) %in% drops)]

write.csv(df,file=paste0(data_in,"airbnb_vienna_raw.csv"))

#drop broken lines - where id is not a character of numbers
df$junk<-grepl("[[:alpha:]]", df$id)
df<-subset(df,df$junk==FALSE)
df$junk <- NULL

#display the class and type of each columns
sapply(df, class)
sapply(df, typeof)


# FORMATTING COLUMNS ------------------------------------------------------

#remove percentage signs
for (perc in c("host_response_rate","host_acceptance_rate")){
  df[[perc]]<-as.numeric(gsub("%","",as.character(df[[perc]])))
}

#remove dollar signs from price variables
for (i in 1:nrow(df)){
  df$price[i] <- as.numeric(gsub("\\$","",as.character(df$price[i])))
}

df$price <- as.numeric(df$price)

#format binary variables
for (binary in c("host_is_superhost","host_has_profile_pic","host_identity_verified", "has_availability", "instant_bookable")){
  df[[binary]][df[[binary]]=="f"] <- 0
  df[[binary]][df[[binary]]=="t"] <- 1
}



# AMENITIES ---------------------------------------------------------------
#amenities
df$amenities<-gsub("\\[","",df$amenities)
df$amenities<-gsub("\\]","",df$amenities)
df$amenities<-gsub('\\"',"",df$amenities)
df$amenities<-as.list(strsplit(df$amenities, ","))

#define levels and dummies 
levs <- levels(factor(unlist(df$amenities)))
df<-cbind(df,as.data.frame(do.call(rbind, lapply(lapply(df$amenities, factor, levs), table))))

drops <- c("amenities","translation missing: en.hosting_amenity_49",
           "translation missing: en.hosting_amenity_50")
df<-df[ , !(names(df) %in% drops)]

# create data frame of the amenities
ams <- df %>% select(-(1:47))

# delete spaces in the beginning and end of the column names, and transfer all to lower case
names(ams) <- gsub(" ","_", tolower(trimws(names(ams))))

# look at the column names we have
levs <- sort(names(ams))


# merge all the columns with the same column name
ams <- as.data.frame(do.call(cbind, by(t(ams), INDICES= names(ams),FUN=colSums)))

# list of key words to merge together
cat <- c( "kitchen", "stove", "oven", "frige","o_machine|ee_machine|coffee", "gril",
         "free.*on_premises", "free.*street", "paid.*on_premis|valet", "paid.*off_premises|self-parking|parking",
         "wifi|internet", "netflix|tv.*netflix", "cable", "tv", "sound_system",
         "toiletries", "shampoo|conditioner", "body_soap|gel", "hair_dryer", "washer", "dryer", "iron",  
         "heating", "air_cond|fan", "balcony|terrace", "garden",
         "onsite_bar|restaurant", "breakfast",  "work|office", "spa",  "fitness|gym",  
         "children|baby|crib|high|corner|chang", "smoking", "housekeeping", "fireplace", "clothing_storage"
         )

# function to merge columns with the same key word in them
for (i in cat) {
  tdf <- ams %>% select(matches(i))

  ams$new_col <- ifelse(rowSums(tdf)>0, 1, 0)
  
  names(ams)[names(ams) == "new_col"] <- paste0("have_", i)
  
  ams <- ams %>% select(-colnames(tdf)) 
  
} 

# keep only columns where the percentage of 1s is at least 1% and at most 99%
selected <- sapply(names(ams), function(x){
  ratio <- sum(ams[[x]])/nrow(ams)*100
  if (between(ratio, 1, 99)) {
    return(TRUE)
  } else { return(FALSE) }
})

amenities <- ams[,selected]

df <- df %>% select((1:47))
  
df <- cbind(df, amenities)

#write csv
write.csv(df,file=paste0(data_out,"airbnb_vienna_cleaned.csv"))
saveRDS(df, paste0(data_out,"airbnb_vienna_cleaned.rds"))
```